---
title: "week 4 & 5"
author: "Siyu Zou"
date: "2023-09-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Control structure
Commonly used control structures are

if and else: testing a condition and acting on it

for: execute a loop a fixed number of times

while: execute a loop while a condition is true

repeat: execute an infinite loop (must break out of it to stop)

break: break the execution of a loop

next: skip an interation of a loop
```{r if else}
x <- runif(n = 1, min = 0, max = 10)
x
y <- if (x > 3) {
    10
} else {
    0
}

y

```

### Q1 use if else
if(<condition1>) {
        ## do something
} else if(<condition2>)  {
        ## do something different
} else {
        ## do something different
}

Let’s use the palmerpenguins dataset and write a if-else statement that

Randomly samples a value from a standard normal distribution (Hint: check out the rnorm(n, mean = 0, sd = 1) function in base R).
If the value is larger than 0, use dplyr functions to keep only the Chinstrap penguins.
Otherwise, keep only the Gentoo penguins.
Re-run the code 10 times and look at output
```{r exercise 1}
library(tidyverse)
library(palmerpenguins)
library( dplyr )
penguins_data <- penguins

i <- rnorm(1,0,1) 

penguins_2 <- penguins_data %>%
  filter(ifelse(i > 0, species == "Chinstrap", species == "Gentoo"))


# Set the seed for reproducibility
set.seed(123)

# Create a loop to run the code 10 times
for (i in 1:10) {
  # Randomly sample a value from a standard normal distribution
  random_value <- rnorm(1, mean = 0, sd = 1)
  
  # Use the value to filter the penguins
  filtered_data <- if (random_value > 0) {
    penguins_data %>% filter(species == "Chinstrap")
  } else {
    penguins_data %>% filter(species == "Gentoo")
  }
  
  # Print the result or perform any other desired operations
  print(filtered_data)
}
```

# for Loops
```{r for}
for (i in 1:10) {
    print(i)
}


## define the loop to iterate over
x <- c("a", "b", "c", "d")

## create for loop
for (i in 1:4) {
    ## Print out each element of 'x'
    print(x[i])
}
```

# seq_along()
The seq_along() function is commonly used in conjunction with for loops in order to generate an integer sequence based on the length of an object (or ncol() of an R object) (in this case, the object x).


The seq_along() function takes in a vector and then returns a sequence of integers that is the same length as the input vector. It doesn’t matter what class the vector is.

Let’s put seq_along() and for loops together.
```{r seq_along}
x <- c("a", "b", "c", "d")
## Generate a sequence based on length of 'x'
for (i in seq_along(x)) {
    print(x[i])
}

# It is not necessary to use an index-type variable (i.e. i).
for (babyshark in x) {
    print(babyshark)
}

# You can use any character index you want (but not with symbols or numbers).
for (RememberToVote in x) {
    print(RememberToVote)
}
```

For one line loops, the curly braces are not strictly necessary.
```{r}
for (i in 1:4) print(x[i])
```

Let’s use the palmerpenguins dataset. Here are the tasks:

Start a for loop
Iterate over the columns of penguins
For each column, extract the values of that column (Hint: check out the pull() function in dplyr).
Using a if-else statement, test whether or not the values in the column are numeric or not (Hint: remember the is.numeric() function to test if a value is numeric).
If they are numeric, compute the column mean. Otherwise, report a NA.
```{r for example}

penguins_data  
col_mean <- vector("double", length(ncol(penguins_data)))
for (i in 1:ncol(penguins_data)){
  col <- pull(penguins_data ,i)  # # Extract the values of the column
  if ( is.numeric(col) ){
    col_mean[i] = mean(col, na.rm = TRUE)
  } else {
    col_mean[i] = NA
  }
  print(mean)
}

```

# Nested for loops
for loops can be nested inside of each other.
```{r}
x <- matrix(1:6, nrow = 2, ncol = 3)
x
for (i in seq_len(nrow(x))) {
    for (j in seq_len(ncol(x))) {
        print(x[i, j])
    }
}
```
Nested loops are commonly needed for multidimensional or hierarchical data structures (e.g. matrices, lists). Be careful with nesting though.

Nesting beyond 2 to 3 levels often makes it difficult to read/understand the code.

If you find yourself in need of a large number of nested loops, you may want to break up the loops by using functions (discussed later).


# while loops
while loops begin by testing a condition.

If it is true, then they execute the loop body.

Once the loop body is executed, the condition is tested again, and so forth, until the condition is false, after which the loop exits.
```{r while}
count <- 0
while (count < 10) {
    print(count)
    count <- count + 1
}

```
while loops can potentially result in infinite loops if not written properly. Use with care!

Sometimes there will be more than one condition in the test.
```{r while example}
z <- 5
set.seed(1)

while (z >= 3 && z <= 10) {
    coin <- rbinom(1, 1, 0.5)

    if (coin == 1) { ## random walk
        z <- z + 1
    } else {
        z <- z - 1
    }
}
print(z)
```

What’s the difference between using one & or two && ?

* If you use only one &, these are vectorized operations, meaning they can return a vector, like this:

-2:2
((-2:2) >= 0) & ((-2:2) <= 0)

* If you use two && (as above), then these conditions are evaluated left to right. For example, in the above code, if z were less than 3, the second test would not have been evaluated.

(2 >= 0) && (-2 <= 0)
(-2 >= 0) && (-2 <= 0)

# repeat loops
repeat initiates an infinite loop right from the start. These are not commonly used in statistical or data analysis applications, but they do have their uses.
```{r repeat}
x0 <- 1
tol <- 1e-8

repeat {
    x1 <- computeEstimate()

    if (abs(x1 - x0) < tol) { ## Close enough?
        break
    } else {
        x0 <- x1
    }
}

```

> The loop above is a bit dangerous because there is no guarantee it will stop.
You could get in a situation where the values of x0 and x1 oscillate back and forth and never converge.
Better to set a hard limit on the number of iterations by using a for loop and then report whether convergence was achieved or not.

# next, break
next is used to skip an iteration of a loop.
```{r next}
for (i in 1:100) {
    if (i <= 20) {
        ## Skip the first 20 iterations
        next
    }
    ## Do something here
}
```
break is used to exit a loop immediately, regardless of what iteration the loop may be on.
```{r break}
for (i in 1:100) {
    print(i)

    if (i > 20) {
        ## Stop loop after 20 iterations
        break
    }
}
```













# function
```{r define function}
my_sum <- function(a, b) {
    a + b
}

## Same but with an extra check to make sure that 'a' and 'b'
## have the same lengths.
my_sum <- function(a, b) {
    ## Check that a and b are of the same length
    stopifnot(length(a) == length(b))
    a + b
}

```

# Loop function
### function loops
lapply(): Loop over a list and evaluate a function on each element

sapply(): Same as lapply but try to simplify the result

apply(): Apply a function over the margins of an array

tapply(): Apply a function over subsets of a vector

mapply(): Multivariate version of lapply (won’t cover)

An auxiliary function split() is also useful, particularly in conjunction with lapply().
```{r loop}



```

### Document your function with roxygen2

```{r roxygen2}
#' Title
#'
#' @param a 
#' @param b 
#'
#' @return
#' @export
#'
#' @examples
my_sum <- function(a, b) {
    ## Check that a and b are of the same length
    stopifnot(length(a) == length(b))
    a + b
}
```

### Test your function with testthat
Below will use two expect_*() functions from testthat. For more details, check https://r-pkgs.org/testing-basics.html.
```{r testhat}
library("testthat")
test_that("my_sum works", {
    x <- seq_len(10)
    expect_equal(my_sum(x, x), x + x)

    expect_error(my_sum(x, seq_len(5)))
})

```

### Share your function in an R package
What even more? Make an R package to share this function.

```{r share function}
install.packages("biocthis")
## Install biocthis if you don't have it
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("biocthis")

## Create an empty R package that is also an
## RStudio project
usethis::create_package("~/Desktop/sum776")

## On the new RStudio window, create the
## scripts that will guide you into making a package
biocthis::use_bioc_pkg_templates()
```

See my example at https://github.com/lcolladotor/sum776 with its documentation website https://lcolladotor.github.io/sum776/. The documentation we wrote for our function is shown at https://lcolladotor.github.io/sum776/reference/my_sum.html.


### Build your own R package with your custom-made ggplot2 theme

## Apply your function
### mapply
```{r mapply}
## Check the arguments to mapply()
args(mapply)

## Apply mapply() to our function my_sum() with the inputs 'x' and 'y'
mapply(sum776::my_sum, x, y)


## Or write an anynymous function that is:
## * not documented
## * not tested
## * not shared
##
## :(
mapply(function(a, b) {
    a + b
}, x, y)

```

### purrr alternative
```{r purrr}
library("purrr") ## part of tidyverse
## Check the arguments of map2_int()
args(purrr::map2_int)

## Apply our function my_sum() to our inputs
purrr::map2_int(x, y, sum776::my_sum)

## You can also use anonymous functions
purrr::map2_int(x, y, function(a, b) {a + b})

## purrr even has a super short formula-like syntax
## where .x is the first input and .y is the second one
purrr::map2_int(x, y, ~ .x + .y)

## This formula syntax has nothing to do with the objects 'x' and 'y'
purrr::map2_int(1:2, 3:4, ~ .x + .y)
```


# Base R loops
## lapply()
it loops over a list, iterating over each element in that list
it applies a function to each element of the list (a function that you specify)
and returns a list (the l in lapply() is for “list”).
It is important to remember that lapply() always returns a list, regardless of the class of the input.
```{r lapply}
lapply
x <- list(a = 1:5, b = rnorm(10))
x

lapply(x, mean)

```

Notice that here we are passing the mean() function as an argument to the lapply() function.
```{r}
purrr::map_dbl(x, mean)
```
What difference do you notice in terms of the output of lapply() and purrr::map_dbl()?

Here is another example of using lapply().
You can use lapply() to evaluate a function multiple times each with a different argument.
```{r lapply 2}
x <- list(a = 1:4, b = rnorm(10), c = rnorm(20, 1), d = rnorm(100, 5))
lapply(x, mean)

```

Next is an example where I call the runif() function (to generate uniformly distributed random variables) four times, each time generating a different number of random numbers.
```{r lapply function}
x <- 1:4
lapply(x, runif)
lapply(x, function(a){a = runif()} )
purrr::map(x, runif)
```

Functions that you pass to lapply() may have other arguments. For example, the runif() function has a min and max argument too.
```{r lapply min max}
x <- 1:4
lapply(x, runif, min = 0, max = 10)
#  Again, this also works with purrr functions.
purrr::map(x, runif, min = 0, max = 10)
```

The lapply() function (and its friends) makes heavy use of anonymous functions
Suppose I wanted to extract the first column of each matrix in the list. I could write an anonymous function for extracting the first column of each matrix.
```{r anonymous}
x <- list(a = matrix(1:4, 2, 2), b = matrix(1:6, 3, 2))
x
lapply(x, function(elt) {
    elt[, 1]
})

#  it’s probably a better idea to define the function separately.
f <- function(elt) {
    elt[, 1]
}
lapply(x, f)
```


## sapply()
The sapply() function behaves similarly to lapply(); the only real difference is in the return value. sapply() will try to simplify the result of lapply() if possible. Essentially, sapply() calls lapply() on its input and then applies the following algorithm:

If the result is a list where every element is length 1, then a vector is returned

If the result is a list where every element is a vector of the same length (> 1), a matrix is returned.

If it can’t figure things out, a list is returned

This can be a source of many headaches and one of the main motivations behind the purrr package. With purrr you know exactly what type of output you are getting!

```{r}
x <- list(a = 1:4, b = rnorm(10), c = rnorm(20, 1), d = rnorm(100, 5))
lapply(x, mean)

# Here’s the result of calling sapply() on the same list.
sapply(x, mean)
```
Because the result of lapply() was a list where each element had length 1, sapply() collapsed the output into a numeric vector, which is often more useful than a list.

With purrr, if I want a list output, I use map(). If I want a double (numeric) output, we can use map_dbl().
```{r purrr map_dbl}
purrr::map(x, mean)   # == lapply

purrr::map_dbl(x, mean)  # == sapply
```

# split()
The split() function takes a vector or other objects and splits it into groups determined by a factor or list of factors.

The arguments to split() are
```{r split}
str(split)
```
where
x is a vector (or list) or data frame
f is a factor (or coerced to one) or a list of factors
drop indicates whether empty factors levels should be dropped

```{r split example}
x <- c(rnorm(10), runif(10), rnorm(10, 1))
f <- gl(3, 10) # generate factor levels
f
split(x, f)
```

A common idiom is split followed by an lapply.
```{r split lpply}
lapply(split(x, f), mean)
```

## Splitting a Data Frame
```{r split data frame}
library("datasets")
head(airquality)

```

We can split the airquality data frame by the Month variable so that we have separate sub-data frames for each month.
```{r}
s <- split(airquality, airquality$Month)
str(s)
```

Then we can take the column means for Ozone, Solar.R, and Wind for each sub-data frame.
```{r}
lapply(s, function(x) {
    colMeans(x[, c("Ozone", "Solar.R", "Wind")])
})
```

Using sapply() might be better here for a more readable output.
```{r}
sapply(s, function(x) {
    colMeans(x[, c("Ozone", "Solar.R", "Wind")])
})

```

Unfortunately, there are NAs in the data so we cannot simply take the means of those variables. However, we can tell the colMeans function to remove the NAs before computing the mean.
```{r remove NA}
sapply(s, function(x) {
    colMeans(x[, c("Ozone", "Solar.R", "Wind")],
        na.rm = TRUE
    )
})
```

We can also do this with purrr as shown below.
```{r purrr example }
purrr::map(s, function(x) {
    colMeans(x[, c("Ozone", "Solar.R", "Wind")], na.rm = TRUE)
})
```

The above is not as condensed as the sapply() output. We can use the superseded (aka no longer supported) function map_dfc():
```{r purrr map_dfc}
purrr::map_dfc(s, function(x) {
    colMeans(x[, c("Ozone", "Solar.R", "Wind")], na.rm = TRUE)
})
```

Or use the currently supported function purrr::list_cbind(). Though we also need to do a bit more work behind the scenes.
```{r}
## Make sure we get data.frame / tibble outputs for each element
## of the list
purrr:::map(s, function(x) {
    tibble::as_tibble(colMeans(x[, c("Ozone", "Solar.R", "Wind")], na.rm = TRUE))
})

## Now we can combine them with list_cbind()
purrr:::map(s, function(x) {
    tibble::as_tibble(colMeans(x[, c("Ozone", "Solar.R", "Wind")], na.rm = TRUE))
}) %>% 
  purrr::list_cbind()
```


## tapply
tapply() is used to apply a function over subsets of a vector. It can be thought of as a combination of split() and sapply() for vectors only. I’ve been told that the “t” in tapply() refers to “table”, but that is unconfirmed.
```{r tapply}
str(tapply)
```

X is a vector
INDEX is a factor or a list of factors (or else they are coerced to factors)
FUN is a function to be applied
… contains other arguments to be passed FUN
simplify, should we simplify the result?

```{r tpply example}
## Simulate some data
x <- c(rnorm(10), runif(10), rnorm(10, 1))
## Define some groups with a factor variable
f <- gl(3, 10)
f
tapply(x, f, mean)
```

We can also apply functions that return more than a single value. In this case, tapply() will not simplify the result and will return a list. Here’s an example of finding the range() (min and max) of each sub-group.
```{r}
tapply(x, f, range)
```

With purrr, we don’t have a tapply() direct equivalent but we can still get similar results thanks to the split() function.

```{r}
split(x, f) %>% 
  purrr::map_dbl(mean)

split(x, f) %>% 
  purrr::map(range)
```


## And we can then add the actual variable it came from with mutate()
```{r}
purrr:::map(s, function(x) {
    tibble::as_tibble(colMeans(x[, c("Ozone", "Solar.R", "Wind")], na.rm = TRUE))
}) %>% purrr::list_cbind() %>% 
    dplyr::mutate(Variable = c("Ozone", "Solar.R", "Wind"))
```

## apply()
The apply() function is used to a evaluate a function (often an anonymous one) over the margins of an array.
 It is most often used to apply a function to the rows or columns of a matrix (which is just a 2-dimensional array). However, it can be used with general arrays, for example, to take the average of an array of matrices. Using apply() is not really faster than writing a loop, but it works in one line and is highly compact.
 
```{r apply}
 str(apply)
```
 X is an array
MARGIN is an integer vector indicating which margins should be “retained”.
FUN is a function to be applied
... is for other arguments to be passed to FUN
 
```{r apply example}
 x <- matrix(rnorm(200), 20, 10)
head(x)

apply(x, 2, mean) ## Take the mean of each column

# I can also compute the sum of each row.
apply(x, 1, sum) ## Take the mean of each row
```

## Col/Row Sums and Means
rowSums = apply(x, 1, sum)
rowMeans = apply(x, 1, mean)
colSums = apply(x, 2, sum)
colMeans = apply(x, 2, mean)

Another nice aspect of these functions is that they are a bit more descriptive. It’s arguably more clear to write colMeans(x) in your code than apply(x, 2, mean).

# Other Ways to Apply
You can do more than take sums and means with the apply() function.
 For example, you can compute quantiles of the rows of a matrix using the quantile() function.
```{r apply quantile}
x <- matrix(rnorm(200), 20, 10)
head(x)
## Get row quantiles
apply(x, 1, quantile, probs = c(0.25, 0.75))

array_branch(x, 1) %>%
    map(quantile, probs = c(0.25, 0.75)) %>%
    map(~ as.data.frame(t(.x))) %>%
    list_rbind()
```
Notice that I had to pass the probs = c(0.25, 0.75) argument to quantile() via the ... argument to apply().


# Vectorizing a Functio
What this means is that we can write function that typically only takes single arguments and create a new function that can take vector arguments.

This is often needed when you want to plot functions.
Here’s an example of a function that computes the sum of squares given some data, a mean parameter and a standard deviation. The formula is 
```{r formula}
sumsq <- function(mu, sigma, x) {
    sum(((x - mu) / sigma)^2)
}

x <- rnorm(100) ## Generate some data
sumsq(mu = 1, sigma = 1, x) ## This works (returns one value)
```
However, passing a vector of mus or sigmas won’t work with this function because it’s not vectorized.
```{r}
sumsq(1:10, 1:10, x) ## This is not what we want
```
There’s even a function in R called Vectorize() that automatically can create a vectorized version of your function.

So we could create a vsumsq() function that is fully vectorized as follows.
```{r}
vsumsq <- Vectorize(sumsq, c("mu", "sigma"))
vsumsq(1:10, 1:10, x)

## The details are a bit complicated though
## as we can see below
vsumsq
```

## Parallelize your functions
We don’t have time to dive into the details, but with furrr we can parallelize purrr functions thanks to the future package. More details at https://furrr.futureverse.org/.

Similarly, with BiocParallel::bplapply() we can parallelize lapply() commands. Details at https://bioconductor.org/packages/release/bioc/vignettes/BiocParallel/inst/doc/Introduction_To_BiocParallel.html and more generally at https://bioconductor.org/packages/BiocParallel/.


# debug

## 3. make it repeatable
Start by creating a reproducible example (*reprex*).
```{r reprex}
library(reprex)
(y <- 1:4)
mean(y)

```

## 4. Figure out where it is

## 5. Fix it and test it
Again, it is very useful to have automated tests in place, which you can implement with *testthat* as showcased in the previous class.

## Somethins is wrong
* message: A generic notification/diagnostic message produced by the message() function; execution of the function continues
* warning: An indication that something is wrong but not necessarily fatal; execution of the function continues. Warnings are generated by the warning() function
* error: An indication that a fatal problem has occurred and execution of the function stops. Errors are produced by the stop() or stopifnot() functions.
* condition: A generic concept for indicating that something unexpected has occurred; programmers can create their own custom conditions if they want.

```{r}
print_message <- function(x) {
    if (x > 0) {
        print("x is greater than zero")
    } else {
        print("x is less than or equal to zero")
    }
    invisible(x)
}

print_message(1)
print_message(NA)
```
What happened?

Well, the first thing the function does is test if x > 0.
But you can’t do that test if x is a NA or NaN value.
R doesn’t know what to do in this case so it stops with a fatal error

We can fix this problem by anticipating the possibility of NA values and checking to see if the input is NA with the is.na() function.
```{R}
print_message2 <- function(x) {
    if (is.na(x)) {
        print("x is a missing value!")
    } else if (x > 0) {
        print("x is greater than zero")
    } else {
        print("x is less than or equal to zero")
    }
    invisible(x)
}

print_message2(NA)
```

## Cases with longer inputs than expected

```{r}
x <- log(c(-1, 2))
print_message2(x)
```
Now what?? Why are we getting this warning?

The warning says “the condition has length > 1 and only the first element will be used”.

The problem here is that I passed print_message2() a vector x that was of length 2 rather then length 1.

Inside the body of print_message2() the expression is.na(x) returns a vector that is tested in the if statement.

However, if cannot take vector arguments, so you get a warning.

The fundamental problem here is that print_message2() is not vectorized.

We can solve this problem two ways.

1. Simply not allow vector arguments.
2. The other way is to vectorize the print_message2() function to allow it to take vector arguments.
```{r}
print_message3 <- function(x) {
    if (length(x) > 1L) {
        stop("'x' has length > 1")
    }
    if (is.na(x)) {
        print("x is a missing value!")
    } else if (x > 0) {
        print("x is greater than zero")
    } else {
        print("x is less than or equal to zero")
    }
    invisible(x)
}
x <- log(c(-1, 2))
print_message3(x)
```
Now when we pass print_message3() a vector, we should get an error.


## Don’t show users the call to help them and you too!
print_message3(1:2)
Error in print_message3(1:2): 'x' has length > 1

I have learned that using that setting *call. = FALSE* when using stop() and/or warning() helps your users by providing them less information that could confuse them.

```{r}
print_message3_no_call <- function(x) {
    if (length(x) > 1L) {
        stop("'x' has length > 1", call. = FALSE)
    }
    if (is.na(x)) {
        print("x is a missing value!")
    } else if (x > 0) {
        print("x is greater than zero")
    } else {
        print("x is less than or equal to zero")
    }
    invisible(x)
}

print_message3_no_call(99:100)
print_message3(99:100)
```
If we compare the error output from print_message3() against print_message3_no_call() we can see that the second scenario doesn’t include information that only we (as the user) have. That’s information that users will copy-paste on their Google searches, which makes finding the right information much harder. As the person trying to help users, knowing how the users called our function is very likely not useful enough information. A reprex is 💯 * ♾️ much better!!!!!


## Errors à la tidyverse
If you want to write error messages similar to those you are used to seeing with tidyverse packages, use rlang. Specifically, switch:

base::stop() with rlang::abort()

base::warning() with rlang::warn()

base::message() with rlang::inform()

```{r rlang}
print_message3_tidyverse <- function(x) {
    if (length(x) > 1L) {
        rlang::abort("'x' has length > 1")
    }
    if (is.na(x)) {
        rlang::warn("x is a missing value!")
    } else if (x > 0) {
        rlang::inform("x is greater than zero")
    } else {
        rlang::inform("x is less than or equal to zero")
    }
    invisible(x)
}
print_message3_tidyverse(99:100)
print_message3_tidyverse(NA)
print_message3_tidyverse(1)
print_message3_tidyverse(-1)
```

Note how rlang by default doesn’t show the function call. The tidyverse style guide has a whole chapter on how to format error messages: https://style.tidyverse.org/error-messages.html. That is how all the R developers in the tidyverse team have been able to provide consistent looking messages to users of all these packages. They also use cli to make their error messages super pretty to read https://rlang.r-lib.org/reference/topic-condition-formatting.html.

In this next example, I:

use cli::cli_abort() instead of rlang::abort()

define len so I can use it in a message with {len}

use {.code something} for inline-markup https://cli.r-lib.org/reference/inline-markup.html

use several of the cli bullets, see https://cli.r-lib.org/reference/cli_bullets.html for the full list

use praise::praise() to get some fun messages to praise our user and lift up their spirits =)

See also praiseMX https://github.com/ComunidadBioInfo/praiseMX which CDSB students built in a course I co-instructed in 2019.
```{r cli}
# install.packages("praiseMx")
print_message3_cli <- function(x) {
    if (length(x) > 1L) {
        len <- length(x)
        
        ## Avoid the print() calls from
        ## https://github.com/ComunidadBioInfo/praiseMX/blob/master/R/praise_crear_emi.R
        praise_mx_log <- capture.output({
            praise_mx <- praiseMX:::praise_bien()
        })
        cli::cli_abort(
            c(
                "This function is not vectorized:",
                "i" = "{.var x} has length {len}.",
                "x" = "{.var x} must have length 1.",
                ">" = "Try using {.code purrr::map(x, print_message3_cli)} to loop your input {.var x} on this function.",
                "v" = praise::praise(),
                "v" = praise_mx
            )
        )
    }
    if (is.na(x)) {
        rlang::warn("x is a missing value!")
    } else if (x > 0) {
        rlang::inform("x is greater than zero")
    } else {
        rlang::inform("x is less than or equal to zero")
    }
    invisible(x)
}

set.seed(20230928)
print_message3_cli(-1:1)
```

# Vectorizing
Vectorizing the function can be accomplished easily with the Vectorize() function.
```{r}
print_message4 <- Vectorize(print_message2)
out <- print_message4(c(-1, 2))
```

# Debugging Tools in R
* traceback(): prints out the function call stack after an error occurs; does nothing if there’s no error
The tidyverse version of this base R function is rlang::last_error() https://rlang.r-lib.org/reference/last_error.html. For base R, you will need to enable tracing with rlang with rlang::global_entrace() https://rlang.r-lib.org/reference/global_entrace.html. It can provide much more user-friendly output.
* debug(): flags a function for “debug” mode which allows you to step through execution of a function one line at a time
* browser(): suspends the execution of a function wherever it is called and puts the function in debug mode
* trace(): allows you to insert debugging code into a function at specific places
* recover(): allows you to modify the error behavior so that you can browse the function call stack

```{r tracebck}
 lm(y ~ x)
# Error in eval(expr, envir, enclos) : object ’y’ not found
traceback()
# 7: eval(expr, envir, enclos)
# 6: eval(predvars, data, env)
# 5: model.frame.default(formula = y ~ x, drop.unused.levels = TRUE)
# 4: model.frame(formula = y ~ x, drop.unused.levels = TRUE)
# 3: eval(expr, envir, enclos)
# 2: eval(mf, parent.frame())
# 1: lm(y ~ x)
```
The traceback() function must be called immediately after an error occurs. Once another function is called, you lose the traceback.

You can see now that the error did not get thrown until the 7th level of the function call stack, in which case the eval() function tried to evaluate the formula y ~ x and realized the object y did not exist.

With rlang, this is how it look:
```{r}
lm(y ~ x)
rlang::last_error()
```

Note that we need to use rlang::global_entrace() first since lm() is a base R function. If we do so, then rlang::last_error() does work.
```{f}
rlang::global_entrace()
lm(y ~ x)
# Error:
# ! object 'y' not found
# Run `rlang::last_trace()` to see where the error occurred.
rlang::last_error()
# <error/rlang_error>
# Error:
# ! object 'y' not found
# ---
# Backtrace:
#     ▆
#  1. └─stats::lm(y ~ x)
#  2.   └─base::eval(mf, parent.frame())
#  3.     └─base::eval(mf, parent.frame())
```


# Error

Note that using message() is better than print() because users can suppress (censor) the messages if they want to with suppressMessages() as shown below. There’s super easy equivalent for print() as capture.output() is more complicated to use.
```{r message}
f <- function() {
    message("This is a message.")
}

f()

suppressMessages(f())
```

## How should errors be handled?
The tryCatch() function is the workhorse of handling errors and warnings in R. The first argument of this function is any R expression, followed by conditions which specify how to handle an error or a warning. The last argument, finally, specifies a function or expression that will be executed after the expression no matter what, even in the event of an error or a warning.
```{r tryCatch}
beera <- function(expr) {
    tryCatch(expr,
        error = function(e) {
            message("An error occurred:\n", e)
        },
        warning = function(w) {
            message("A warning occured:\n", w)
        },
        finally = {
            message("Finally done!")
        }
    )
}
```

Let's try it
```{r}
beera({
    2 + 2
})

beera({
    "two" + 2
})

beera({
    as.numeric(c(1, "two", 3))
})
```
Notice that we’ve effectively transformed errors and warnings into messages.

Now that you know the basics of generating and catching errors you’ll need to decide when your program should generate an error. My advice to you is to limit the number of errors your program generates as much as possible. Even if you design your program so that it’s able to catch and handle errors, the error handling process slows down your program by orders of magnitude. Imagine you wanted to write a simple function that checks if an argument is an even number. You might write the following:
```{r}
is_even <- function(n) {
    n %% 2 == 0
}

is_even(768)
is_even("two")
```

You can see that providing a string causes this function to raise an error. You could imagine though that you want to use this function across a list of different data types, and you only want to know which elements of that list are even numbers. You might think to write the following:
```{r}
is_even_error <- function(n) {
    tryCatch(n %% 2 == 0,
        error = function(e) {
            FALSE
        }
    )
}

is_even_error(714)
is_even_error("eight")
```


```{r}
is_even_check <- function(n) {
    is.numeric(n) && n %% 2 == 0
}

is_even_check(1877)

is_even_check("twelve")
```
Notice that by using is.numeric() before the “AND” operator (&&), the expression n %% 2 == 0 is never evaluated. This is a programming language design feature called “short circuiting.” The expression can never evaluate to TRUE if the left hand side of && evaluates to FALSE, so the right hand side is ignored.


To demonstrate the difference in the speed of the code, we will use the microbenchmark package to measure how long it takes for each function to be applied to the same data.
```{r}
# install.packages("microbenchmark")
library(microbenchmark)
microbenchmark(sapply(letters, is_even_check))
microbenchmark(sapply(letters, is_even_error))
```

