---
title: "week 2"
author: "Siyu Zou"
date: "2023-09-05"
output: html_document
---

```{r literature, include=FALSE}
citation("rmarkdown")
citation("tidyverse") 
```

```{r week 2 reading and writing data}
getwd()
# setwd()

install.packages('here')
library(here)
library(sessioninfo)
# install.packages("palmerpenguins")
library("palmerpenguins")
```

```{r finding files locally}
here::here()
list.files(here::here())
# here makes the sharing code more easily, you don't need to change your dictionary

# file.exists(here("my", "relative", "path")): logical test if the file exists
# dir.create(here("my", "relative", "path")): create a folder
# list.files(here("my", "relative", "path")): list contents of folder
# file.create(here("my", "relative", "path")): create a file
# file.remove(here("my", "relative", "path")): delete a file
```

```{r txt or csv}
# base R
# read.csv() or read.table()
# write.table() or writeLines()
# df <- read.csv(here("data", "team_standings.csv"))

# reading data files with read.table()

library(readr) # read_table() and read_csv()
# teams <- read_csv(here("data", "team_standings.csv")) 
# read_* tries to guess the class of every column, and much faster than base R (read.*)

```

```{r source}
# read function from other R script file into R
source(here::here("functions.R"))  
# you don't need to open the file and read those functions 
# cons: make the code shorter
```

```{r saving}
# .RData: Stores multiple R objects
# .Rda: This is short for .RData and is equivalent.
# .Rds: Stores a single R object

a <- september
b <- tuesday
save(a,b, file = here(".RData"))
saveRDS(a,file = here("a.Rds"))

load(here("a.Rds"))
```

```{r tidyverse}
library(tidyverse)
# tibble is a nice way to show data than data.frame, more informative

chicago <- readRDS(here("data", "chicago.rds"))  # this is a data.frame

dim(chicago)  

str(chicago)  # data structure 

#  convert this data.frame to a tibble you would use the following:
str(as_tibble(chicago))

# create a tibble
tibble(
    a = 1:5,
    b = 6:10,
    c = 1,
    z = (a + b)^2 + c
)

# Tibbles can have column names that are not allowed in data.frame
tibble(
    `two words` = 1:5,
    `12` = "numeric",
    `:)` = "smile",
)

```
[[ can subset by name or position
$ only subsets by name
```{r seusetting}
df <- tibble(
    a = 1:5,
    b = 6:10,
    c = 1,
    z = (a + b)^2 + c
)

# Extract by name using $ or [[]]
df$z
df[["z"]]
# Extract by position requires [[]]
df[[4]]

```
# Dplyr
mutate(): add new variables/columns or transform existing variables
select(): return a subset of the columns of a data frame, using a flexible notation
filter(): extract a subset of rows from a data frame based on logical conditions
summarise() / summarize(): generate summary statistics of different variables in the data frame, possibly within strata
## %>%: the “pipe” operator is used to connect multiple verb actions together into a pipeline

```{r dplyr}
library(dplyr)

chicago <- as_tibble(chicago) # dataset
str(chicago)
```

# select
```{r select}
# subset
chicage %>%
  select(city:dptp)
# keep every variable that ends with"2"
subset <- select(chicago, ends_with("2"))
str(subset)
```

# filter
```{r filter}
chicago_filter <- chicago %>%
  filter(pm25tmean2 > 30)

summary(chicago_filter$pm25tmean2)
```

# logical operations
%in%	(Included in) 
example: 	city %in% c("chic", "bmore")

is.na()	Is a missing value	
example: is.na(pm10tmean2)

# arrange
 reorder rows of a data frame according to one of the variables/columns.
```{r arrange}
chicago <- arrange(chicago, date)

# descending order 
chicago <- arrange(chicago, desc(date))
```

# Renaming a variable
```{r rename}
chicago <- rename(chicago, dewpoint = dptp, pm25 = pm25tmean2)
head(chicago[, 1:5], 3)
```

# mutate
create new variables drives from existing variables
```{r mutate}
chicago <- mutate(chicago, pm25detrend = pm25 - mean(pm25, na.rm = TRUE))
head(chicago)

# transmute() function, which does the same thing as mutate() but then drops all non-transformed variables.
head(transmute(chicago,
    pm10detrend = pm10tmean2 - mean(pm10tmean2, na.rm = TRUE),
    o3detrend = o3tmean2 - mean(o3tmean2, na.rm = TRUE)  # ignore the missing 
))
# Note that there are only two columns in the transmuted data frame.
```

# group_by
generate summary statistics from the data frame within strata defined by a variable
1. Splitting a data frame into separate pieces defined by a variable or group of variables (group_by())
2. Then, applying a summary function across those subsets (summarize())
```{r group_by}
chicago <- mutate(chicago, year = as.POSIXlt(date)$year + 1900)
years <- group_by(chicago, year)
summarize(years,
    pm25 = mean(pm25, na.rm = TRUE),
    o3 = max(o3tmean2, na.rm = TRUE),
    no2 = median(no2tmean2, na.rm = TRUE)
)
```

# %>%
## The pipeline operator %>% is very handy for stringing together multiple dplyr functions in a sequence of operations
hard to read: third(second(first(x)))
```{r pipe}
chicago %>%
    mutate(year = as.POSIXlt(date)$year + 1900) %>%
    group_by(year) %>%
    summarize(
        pm25 = mean(pm25, na.rm = TRUE),
        o3 = max(o3tmean2, na.rm = TRUE),
        no2 = median(no2tmean2, na.rm = TRUE)
    )

# another example: by month
mutate(chicago, month = as.POSIXlt(date)$mon + 1) %>%
    group_by(month) %>%
    summarize(
        pm25 = mean(pm25, na.rm = TRUE),
        o3 = max(o3tmean2, na.rm = TRUE),
        no2 = median(no2tmean2, na.rm = TRUE)
    )

```

# slice_*()
 allow you to see a sample of random rows in random order.
## slice_head() or slice_tail() to take a look at the top rows or bottom rows of your tibble.
```{r slice_sample}
slice_sample(chicago, n = 10)
# show the first 5 rows
slice_head(chicago, n = 5)
# show the last 5 rows
slice_tail(chicago, n = 5)
```




# Tidy data

```{r tidyr}
library(tidyr)
relig_income

library(tidyverse)

# 转换 Pivot data from wide to long
relig_income %>%
    pivot_longer(-religion, names_to = "income", values_to = "respondents") %>%
    mutate(religion = factor(religion), income = factor(income))

```

# A subset of the “Tidyverse” packages include:

ggplot2: a plotting system based on the grammar of graphics

magrittr: defines the %>% operator for chaining functions together in a series of operations on data

dplyr: a suite of (fast) functions for working with data frames

tidyr: easily tidy data with pivot_wider() and pivot_longer() functions (also separate() and unite())

```{r start}

library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
```

```{r pivot_longer}
# install.packages("gapminder")
library(gapminder)
gapminder

```

# separate() and unite()
unite(): combine contents of two or more columns into a single column
separate(): separate contents of a column into two or more columns
```{r separate}

gapminder %>%
    unite(
        col = "country_continent_year",
        country:year,
        sep = "_"
    )

# separate
gapminder %>%
    unite(
        col = "country_continent_year",
        country:year,
        sep = "_"
    ) %>%
    separate(
        col = "country_continent_year",
        into = c("country", "continent", "year"),
        sep = "_"
    )
```


# Joining data in R
here are three families of verbs designed to work with relational data:

## Mutating joins: A mutating join allows you to combine variables from two tables. 
## Filtering joins: Filtering joins match observations in the same way as mutating joins, but affect the observations, not the variables (i.e. filter observations from one data frame based on whether or not they match an observation in the other).
Two types: semi_join(x, y) and anti_join(x, y).
## Set operations: Treat observations as if they were set elements


# mutating join
left_join() is useful for merging a “large” data frame with a “smaller” one while retaining all the rows of the “large” data frame

inner_join() gives you the intersection of the rows between two data frames

right_join() is like left_join() with the arguments reversed (likely only useful at the end of a pipeline)
```{r mutating join}
library(tidyverse)

outcomes <- tibble(
    id = rep(c("a", "b", "c"), each = 3),
    visit = rep(0:2, 3),
    outcome = rnorm(3 * 3, 3)
)

print(outcomes)

# second table
subjects <- tibble(
    id = c("a", "b", "c"),
    house = c("detached", "rowhouse", "rowhouse")
)

print(subjects)
```


# left join
The bottom line for left_join() is that it always retains the values in the “left” argument (in this case the outcomes table).

If there are no corresponding values in the “right” argument, NA values will be filled in.
```{r left join}
left_join(x = outcomes, y = subjects, by = "id")
```

# inner join
The inner_join() function only retains the rows of both tables that have corresponding values. 
```{r inner join}
inner_join(x = outcomes, y = subjects, by = c("id", "visit"))
```

# right join
The right_join() function is like the left_join() function except that it gives priority to the “right” hand argument.
```{r right join}
right_join(x = outcomes, y = subjects, by = c("id", "visit"))
```

```{r end}
session_info()

```

