---
title: "Week 3"
author: "Siyu Zou"
date: "2023-09-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
```

## ggplot2
A 'data frame': stores all of the data that will be displayed on the plot

aesthetic mappings: describe how data are mapped to color, size, shape, location

geoms: geometric objects like points, lines, shapes

facets: describes how conditional/panel plots should be constructed

stats: statistical transformations like binning, quantiles, smoothing

scales: what scale an aesthetic map uses (example: left-handed = red, right-handed = blue)

coordinate system: describes the system in which the locations of the geoms will be drawn



```{r data, echo=FALSE}
maacs <- read_csv(here("data", "bmi_pm25_no2_sim.csv"),
    col_types = "nnci"
)
maacs

```

```{r ggplot}
g <- ggplot(maacs, aes(
    x = logpm25,
    y = NocturnalSympt
))
summary(g)

class(g)
```

normally if you were to print() a ggplot object a plot would appear on the plot device, however, our object g actually does not contain enough information to make a plot yet.
```{r only g}
g <- maacs %>%
    ggplot(aes(logpm25, NocturnalSympt))
print(g)
```


To make a scatter plot, we need add at least one geom, such as points.
```{r use geom}
g <- maacs %>%
    ggplot(aes(logpm25, NocturnalSympt))
g + geom_point()

```

## ading more layers
```{r smooth}
g +
    geom_point() +
    geom_smooth()
```

The default smoother is a loess smoother, which is flexible and nonparametric but might be too flexible for our purposes. Perhaps we’d prefer a simple linear regression line to highlight any first order trends. We can do this by specifying method = "lm" to geom_smooth().

```{r smooth method = lm}
g +
    geom_point() +
    geom_smooth(method = "lm")
```


## Exercise
```{r question 1}
# try it yourself

library("palmerpenguins")
penguins


penguins %>%
  ggplot(aes(x = flipper_length_mm, y = bill_length_mm, color = species )) +  # global environment, apply to all layer
  geom_point( ) +
  geom_smooth(method = "lm" ) 


penguins %>%
  ggplot(aes(x = flipper_length_mm, y = bill_length_mm )) +  # global environment
  geom_point( aes(color = species )) +  # only apply to this layer
  geom_smooth(method = "lm" ) 
```

## facets
stratify the scatter plot of PM2.5 and nocturnal symptoms by the BMI category (bmicat) variable, which indicates whether an individual is overweight or not.
```{r facet_grid}
# We want one row and two columns, one column for each weight category. So we specify bmicat on the right hand side of the forumla passed to facet_grid().
g +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_grid(. ~ bmicat)


```
This plot suggests that overweight individuals may be more susceptible to the effects of PM2.5.

## Modifying geom properties
```{r map aesthetics to constants}
g + 
  geom_point(color = "steelblue", size = 4, alpha = 1 / 2)
```
For example, here we modify the points in the scatterplot to make the color “steelblue”, the size larger, and the alpha transparency greater.


```{r map aesthetics to variables}
g + 
  geom_point(aes(color = bmicat), size = 4, alpha = 1 / 2)

```
For example, we can map the aesthetic color to the variable bmicat, so the points will be colored according to the levels of bmicat.
We use the aes() function to indicate this difference from the plot above.

## Customizing the smooth
For example, we can customize the smoother that we overlay on the points with geom_smooth().
Here we change the line type and increase the size from the default. We also remove the shaded standard error from the line.
```{r customize aspects of geoms}
g +
    geom_point(aes(color = bmicat),
        size = 2,
        alpha = 1 / 10  # transparancy greater
    ) +
    geom_smooth(
        linewidth = 4,
        linetype = 3,
        method = "lm",
        se = FALSE
    )
```

## Changing the theme
The default theme for ggplot2 uses the 'gray background' with white grid lines.
For things that only make sense globally, use theme(), i.e. theme(legend.position = "none"). 

theme_gray(): The default theme (gray background)
theme_bw(): More stark/plain
```{r theme change}
g +
    geom_point(aes(color = bmicat)) +
    theme_bw(base_family = "Times")

g +
    geom_point(aes(color = bmicat)) +
    theme_bw(base_size = 20) # make everything bigger

g +
    geom_point(aes(color = bmicat)) +
    theme_dark(base_size = 20)

g +
    geom_point(aes(color = bmicat)) +
    theme_classic(base_size = 20) + 
  theme(legend.position = "bottom")

```


## Modifing lables
xlab() for x-axis labels
ylab() for y-axis labels
ggtitle() for specifying plot titles
```{r labels}
g +
    geom_point(aes(color = bmicat)) +
    labs(title = "MAACS Cohort") +
    labs(
        x = expression("log " * PM[2.5]),
        y = "Nocturnal Symptoms",
        caption = "biostat776"  # 图片说明
    )

```

### axis limits
If you make a lot of time series plots, you often want to restrict the range of the y-axis while still plotting all the data.
```{r axis limits}
testdat <- data.frame(
    x = 1:100,
    y = rnorm(100)
)
testdat[50, 2] <- 100 ## Outlier!
plot(testdat$x,
    testdat$y,
    type = "l",
    ylim = c(-3, 3)
)
```
Here, we have restricted the y-axis range to be between -3 and 3, even though there is a clear outlier in the data.

```{r default}
g <- ggplot(testdat, aes(x = x, y = y))
g + geom_line()

```

One might think that modifying the ylim() attribute would give you the same thing as the base plot, but it doesn’t (?????)
```{r ylim}
g +
    geom_line() +
    ylim(-3, 3)

```
Effectively, what this does is subset the data so that only observations between -3 and 3 are included, then plot the data.

To plot the data without subsetting it first and still get the restricted range, you have to do the following.
```{r limit}
g +
    geom_line() +
    coord_cartesian(ylim = c(-3, 3))
```

# BBC's plots
```{r bbplot}
## Install bbplot
# remotes::install_github("bbc/bbplot")
```

```{r bbc_style}
## Basic ggplot2 object with our data
g <- maacs %>%
    ggplot(aes(logpm25, NocturnalSympt))

## A plot we made before, but this time without the SE lines
g +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE) +
    facet_grid(. ~ bmicat)
```


```{r bbc}
## Now let's add bbplot::bbc_style()
g +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE) +
    facet_grid(. ~ bmicat) +
    bbplot::bbc_style()

```

Next, we can use the blue color that the BBC frequently uses and further improve our plot by adding a title and subtitle.

```{r blue color}
g +
    geom_point() +
    geom_smooth(colour = "#1380A1", method = "lm", se = FALSE) +
    facet_grid(. ~ bmicat) +
    bbplot::bbc_style() +
    labs(
        title = "Child asthma's link to air quality worsens in overweight children",
        subtitle = "Number of days with symptoms vs PM2.5 by weight group"
    )

```

```{r drop points}
g +
    geom_smooth(aes(colour = bmicat), method = "lm", se = FALSE, linewidth = 2) +
    scale_colour_manual(values = c("#FAAB18", "#1380A1")) +
    bbplot::bbc_style() +
    labs(
        title = "Child asthma's link to air quality worsens in overweight children",
        subtitle = "Number of days with symptoms vs PM2.5 by weight group"
    )
```

# What about Barbenheimer?
```{r}
## Install ThemePark from GitHub
# remotes::install_github("MatthewBJane/theme_park")
```

```{r}
## Barbie-inspired theme
g +
    geom_smooth(aes(colour = bmicat), method = "lm", se = FALSE, linewidth = 2) +
    scale_colour_manual(values = c("#FAAB18", "#1380A1")) +
    ThemePark::theme_barbie() +
    labs(
        title = "Child asthma's link to air quality worsens in overweight children",
        subtitle = "Number of days with symptoms vs PM2.5 by weight group"
    )

## Oppenheimer-inspired theme
g +
    geom_smooth(aes(colour = bmicat), method = "lm", se = FALSE, linewidth = 2) +
    scale_colour_manual(values = c("#FAAB18", "#1380A1")) +
    ThemePark::theme_oppenheimer() +
    labs(
        title = "Child asthma's link to air quality worsens in overweight children",
        subtitle = "Number of days with symptoms vs PM2.5 by weight group"
    )

```

Nowadays you can find many ggplot2 themes online. Just search your favorite movie / keyword + ggplot2 on Google Search or your favorite browser search tool.

Note that one great package with many themes is the ggthemes package showcased at https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/.

#### Install ggthemes from CRAN
```{r}
# install.packages("ggthemes")

## Your favorite statistics class theme ;)
## I bet that you could fool a few people into thinking
## that you are not using R ^_^'
g +
    geom_smooth(aes(colour = bmicat), method = "lm", se = FALSE, linewidth = 2) +
    scale_colour_manual(values = c("#FAAB18", "#1380A1")) +
    ggthemes::theme_stata() +
    labs(
        title = "Child asthma's link to air quality worsens in overweight children",
        subtitle = "Number of days with symptoms vs PM2.5 by weight group"
    )
```

# Can we… make them interactive?
Yes we can! Well, not always, but you can always try. Simply save your plot as an object, then feed that object into plotly::ggplotly(). plotly is available from CRAN as is documented at https://plotly.com/r/. That’s how I made a lot of the interactive plots in spatialLIBD that you can see at http://spatial.libd.org/spatialLIBD/.
```{r interactive}
## Save our plot into an object
g_complete <- g +
    geom_point(aes(colour = bmicat)) +
    geom_smooth(aes(colour = bmicat), method = "lm", se = FALSE, linewidth = 2) +
    scale_colour_manual(values = c("#FAAB18", "#1380A1"))

## Make it interactive with plotly::ggplotly()
plotly::ggplotly((g_complete))

```



