---
title: "Week 7"
author: "Siyu Zou"
date: "2023-10-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(forcats)
library(tidyverse)
```

# factor
At the end of this lesson you will:

How to create factors and some challenges working with them in base R
An introduction to the *forcats* package in the *tidyverse* to work with categorical variables in R

Factors are used for working with categorical variables, or variables that have a fixed and known set of possible values (income bracket, U.S. state, political affiliation).

Factors are useful when:

* You want to include categorical variables in regression models
* You want to plot categorical data (e.g. want to map categorical variables to aesthetic attributes)
* You want to display character vectors in a non-alphabetical order



## Factor basics
You can fix both of these problems with a factor.

To create a factor you must start by creating a list of the valid levels:
```{r factor creat a level}
month_levels <- c(
    "Jan", "Feb", "Mar", "Apr", "May", "Jun",
    "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"
)
```
Now we can create a factor with the factor() function defining the levels argument:
```{r factor add level}
x <- c("Dec", "Apr", "Jan", "Mar")
sort(x)
y <- factor(x, levels = month_levels)
y
```
We can see what happens if we try to sort the factor:
```{r}
sort(y)
```
We can also check the attributes of the factor:
```{r}
attributes(y)
```

If you want to access the set of levels directly, you can do so with levels():
```{r}
levels(y)
```

Any values not in the level will be silently converted to NA:
```{r typo NA}
x_typo <- c("DEC", "dec", "jan")
y_typo <- factor(x_typo, levels = month_levels)
y_typo
```

## Challenges working with categorical data
```{r}
library(tidyverse)

x1_original <- c(10, 10, 10, 50, 60, 20, 20, 40)
x1_factor <- factor(x1_original)
attributes(x1_factor)
```

```{r tibble}
tibble(x1_original, x1_factor) %>%
    mutate(x1_numeric = as.numeric(x1_factor))
```
Instead of creating a new variable with a numeric version of the value of the factor variable x1_factor, the variable loses the original numerical categories and creates a factor number (i.e., 10 is mapped to 1, 20 is mapped to 2, and 40 is mapped to 3, etc).

This result is unexpected because base::as.numeric() is intended to recover numeric information by coercing a character variable.




The tidyverse is part of this movement, with functions from the readr package defaulting to leaving strings as-is. Others used to chose to add options(stringAsFactors = FALSE) into their start up commands to override R’s default of stringsAsFactors = TRUE in functions such as read.table(). However, that is no longer needed in recent versions of R as the default has become stringsAsFactors = FALSE as documented on the official R blog: https://blog.r-project.org/2020/02/16/stringsasfactors/.




## Factors when modeling data
Factors are still necessary for some data analytic tasks. The most salient case is in statistical modeling.
When you pass a factor variable into lm() or glm(), R automatically creates indicator (or more colloquially *‘dummy’)* variables for each of the levels and picks one as a reference group.

For simple cases, this behavior can also be achieved with a character vector.

However, to choose *which level to use as a reference level* or to order classes, factors must be used.

Consider a vector of character strings with three income levels:
```{r}
income_level <- c(
    rep("low", 10),
    rep("medium", 10),
    rep("high", 10)
)
income_level
```

Here, it might make sense to use the lowest income level (low) as the reference class so that all the other coefficients can be interpreted in comparison to it.

However, R would use high as the reference by default because ‘h’ comes before ‘l’ in the alphabet.

```{r}
x <- factor(income_level)
x
```

```{r}
y <- rnorm(30) # generate some random obs from a normal dist
lm(y ~ x)  # high is the ref

x <- relevel(x, "low")
lm(y ~ x)  # low is the ref
```


### Memory req for factors and character strings

In early versions of R, storing categorical data as a factor variable was considerably more efficient than storing the same data as strings, because factor variables only store the factor labels once.
However, R now uses a global string pool, so each unique string is only stored once, which means storage is now less of an issue.
```{r}
income_level <- c(
    rep("low", 10000),
    rep("medium", 10000),
    rep("high", 10000)
)

format(object.size(income_level), units = "Kb") # size of the character string

format(object.size(factor(income_level)), units = "Kb") # size of the factor

```



## forcats
Next, we will introduce the forcats package, which is part of the core tidyverse, but can also be loaded directly

library("forcats")

It provides tools for dealing with *cat*egorical variables (and it’s an anagram of factors!) using a wide range of helpers for working with factors.

### Example: 
For the rest of this lecture, we are going to use the gss_cat dataset that is installed when you load forcats.
```{r}
gss_cat
```

When factors are stored in a tibble, you cannot see their levels so easily. One way to view them is with count():
```{r}
gss_cat %>%
    count(race)

gss_cat %>%
    ggplot(aes(x = race)) +
    geom_bar()
```
When working with factors, the two most common operations are

Changing the order of the levels
Changing the values of the levels


### Modifying factor order
It’s often useful to change the order of the factor levels in a visualization.

Let’s explore the *relig* (religion) factor
```{r modify order}
gss_cat %>%
    count(relig)

attributes(gss_cat$relig)
```

The first level is “No answer” followed by “Don’t know”, and so on.

Imagine you want to explore the average number of hours spent watching TV (*tvhours*) per day across religions (*relig*):
```{r}
relig_summary <- gss_cat %>%
    group_by(relig) %>%
    summarise(
        tvhours = mean(tvhours, na.rm = TRUE),
        n = n()
    )

relig_summary %>%
    ggplot(aes(x = tvhours, y = relig)) +
    geom_point()
```
The y-axis lists the levels of the relig factor in the order of the levels.

However, it is hard to read this plot because there’s no overall pattern.


## fct_reorder
We can improve it by reordering the levels of relig using fct_reorder(). 
fct_reorder(.f, .x, .fun) takes three arguments:
* .f, the factor whose levels you want to modify.
* .x, a numeric vector that you want to use to reorder the levels.
* Optionally, .fun, a function that’s used if there are multiple values of x for each value of f. The default value is median.

```{r}
relig_summary %>%
    ggplot(aes(
        x = tvhours,
        y = fct_reorder(.f = relig, .x = tvhours)
    )) +
    geom_point()
```
Reordering religion makes it much easier to see that people in the “Don’t know” category watch much more TV, and Hinduism & Other Eastern religions watch much less.

As you start making more complicated transformations, I recommend moving them out of aes() and into a separate mutate() step.


Example
You could rewrite the plot above as:
```{r}
relig_summary %>%
    mutate(relig = fct_reorder(relig, tvhours)) %>%
    ggplot(aes(x = tvhours, y = relig)) +
    geom_point()
```

Another example
What if we create a similar plot looking at how average age varies across reported income level?
```{r}
rincome_summary <-
    gss_cat %>%
    group_by(rincome) %>%
    summarise(
        age = mean(age, na.rm = TRUE),
        n = n()
    )

## Original rincome order
rincome_summary %>%
    ggplot(aes(x = age, y = rincome)) +
    geom_point()
```

## rincome re-ordered by age's values
```{r}
rincome_summary %>%
    ggplot(aes(x = age, y = fct_reorder(.f = rincome, .x = age))) +
    geom_point()

```

Reserve fct_reorder() for factors whose levels are arbitrarily ordered.


#### Question
Let’s practice fct_reorder(). Using the palmerpenguins dataset,

Calculate the average bill_length_mm for each species
Create a scatter plot showing the average for each species.
Go back and reorder the factor species based on the average bill length from largest to smallest.
Now order it from smallest to largest
```{r}
library(palmerpenguins)
penguins

bill_length_summary <- penguins %>%
  group_by(species) %>%
  summarise(
    mean_bill_length_mm = mean(bill_length_mm, na.rm = TRUE),
    n = n()
  )

bill_length_summary %>%
  mutate(species = fct_reorder(species,mean_bill_length_mm ) ) %>%
  ggplot(aes(x = mean_bill_length_mm, y = species ))  + 
    geom_point()
  
```

### fct_relevel
However, it does make sense to pull “Not applicable” to the front with the other special levels.

You can use fct_relevel().

It takes a factor, f, and then any number of levels that you want to move to the front of the line.
```{r}
rincome_summary %>%
    ggplot(aes(age, fct_relevel(rincome, "Not applicable"))) +
    geom_point()
```

Any levels not mentioned in fct_relevel will be left in their existing order.

Another type of reordering is useful when you are coloring the lines on a plot. *fct_reorder2(f, x, y)* reorders the factor f by the y values associated with the largest x values.
```{r}
by_age <-
    gss_cat %>%
    filter(!is.na(age)) %>%
    count(age, marital) %>%
    group_by(age) %>%
    mutate(prop = n / sum(n))

by_age %>%
    ggplot(aes(age, prop, colour = marital)) +
    geom_line(na.rm = TRUE)

by_age %>%
    ggplot(aes(age, prop, colour = fct_reorder2(marital, age, prop))) +
    geom_line() +
    labs(colour = "marital")

```


### fct_infreq
Finally, for bar plots, you can use fct_infreq() to order levels in decreasing frequency: this is the simplest type of reordering because it doesn’t need any extra variables. Combine it with *fct_rev()* if you want them in increasing frequency so that in the bar plot largest values are on the right, not the left.
```{r order in frequency}
gss_cat %>%
    mutate(marital = marital %>% fct_infreq() %>% fct_rev()) %>%
    ggplot(aes(marital)) +
    geom_bar()
```

```{r}
fct_rev(x$species)
# same to this: 
levels(x$species) <- rev(x$species)
```

## Modifying factor levels
#### fct_recode
The most general and powerful tool is fct_recode(). It allows you to recode, or change, the value of each level. For example, take the gss_cat$partyid:
```{r fct_recode}
gss_cat %>%
    count(partyid)
```

The levels are terse and inconsistent.

Let’s tweak them to be longer and use a parallel construction.

Like most rename and recoding functions in the tidyverse:

* *the new values go on the left*
* the old values go on the right

```{r}
gss_cat %>%
    mutate(partyid = fct_recode(partyid,
        "Republican, strong"    = "Strong republican",
        "Republican, weak"      = "Not str republican",
        "Independent, near rep" = "Ind,near rep",
        "Independent, near dem" = "Ind,near dem",
        "Democrat, weak"        = "Not str democrat",
        "Democrat, strong"      = "Strong democrat"
    )) %>%
    count(partyid)


```

fct_recode() will leave the levels that aren’t explicitly mentioned as is, and will warn you if you accidentally refer to a level that doesn’t exist.

#### combine 
To combine groups, you can assign multiple old levels to the same new level:
```{r}
gss_cat %>%
    mutate(partyid = fct_recode(partyid,
        "Republican, strong"    = "Strong republican",
        "Republican, weak"      = "Not str republican",
        "Independent, near rep" = "Ind,near rep",
        "Independent, near dem" = "Ind,near dem",
        "Democrat, weak"        = "Not str democrat",
        "Democrat, strong"      = "Strong democrat",
        "Other"                 = "No answer",
        "Other"                 = "Don't know",
        "Other"                 = "Other party"
    )) %>%
    count(partyid)
```


### fct_collapse
If you want to collapse a lot of levels, fct_collapse() is a useful variant of fct_recode().
For each new variable, you can provide a vector of old levels:
```{R}
gss_cat %>%
    mutate(partyid = fct_collapse(partyid,
        "other" = c("No answer", "Don't know", "Other party"),
        "rep" = c("Strong republican", "Not str republican"),
        "ind" = c("Ind,near rep", "Independent", "Ind,near dem"),
        "dem" = c("Not str democrat", "Strong democrat")
    )) %>%
    count(partyid)
```

#### fct_lump_*
Sometimes you just want to lump together the small groups to make a plot or table simpler.

That’s the job of the fct_lump_*() family of functions.

fct_lump_lowfreq() is a simple starting point that progressively lumps the smallest groups categories into “Other”, always keeping “Other” as the smallest category.
```{r}
gss_cat %>%
    mutate(relig = fct_lump_lowfreq(relig)) %>%
    count(relig)
```

In this case it’s not very helpful: it is true that the majority of Americans in this survey are Protestant, but we’d probably like to see some more details!
Instead, we can use the *fct_lump_n()* to specify that we want exactly 10 groups:

```{r}
gss_cat %>%
    mutate(relig = fct_lump_n(relig, n = 10)) %>%
    count(relig, sort = TRUE) %>%
    print(n = Inf)
```
Read the documentation to learn about fct_lump_min() and fct_lump_prop() which are useful in other cases.


## Ordered factors
There’s a special type of factor that needs to be mentioned briefly: ordered factors.
Ordered factors, created with ordered(), imply a strict ordering and equal distance between levels:

The first level is “less than” the second level by the same amount that the second level is “less than” the third level, and so on…
```{r}
ordered(c("a", "b", "c"))

factor(c("a","b","c"))

model.matrix(~factor(c("a","b","c")))
model.matrix(~ordered(c("a","b","c")))
```
However, in practice, ordered() factors behave very similarly to regular factors.




















