---
title: "week 6"
author: "Siyu Zou"
date: "2023-10-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Working with dates and times
Recognize the *Date*, *POSIXct* and *POSIXlt* class types in R to represent dates and times
Learn how to create date and time objects in R using functions from the lubridate package
Dates are stored internally as the number of days since 1970-01-01
Times are stored internally as the number of seconds since 1970-01-01
## lubridate package
lubridate is installed when you install tidyverse, but it is not loaded when you load tidyverse
You should always use the simplest possible data type that works for your needs.
That means if you can use a date instead of a date-time, you should.
```{r time}
library("tidyverse")
library("lubridate")

today() 
now()
base::Sys.Date()
base::Sys.time()
```
Typical ways to create a date/time in R
From a string
From individual date-time components
From an existing date/time object

### 1.From a string
```{r}
x <- today()
class(x)
ymd("1970-01-01")
mdy("January 31st, 2017")
dmy("31-Jan-2017")

## Base R versions
as.Date("1970-01-01")
```
 Quickly becomes more complicated.
 We quickly need to get familiarized with formats for specifying dates, and that's complicated.
```{r}
?strptime
as.Date("January 31st, 2017", "%B %dst, %Y")
as.Date(gsub("st,", "", "January 31st, 2017"), "%B %d %Y")
```

When reading in data with read_csv(), you may need to read in as character first and then convert to date/time
Date objects have their own special print() methods that will always format as “YYYY-MM-DD”
These functions also take unquoted numbers.
```{r unquoted}
ymd(20170131)
```

## Alternate Formulations
Different locales have different ways of formatting dates
```{r different formats}
ymd("2016-09-13") ## International standard
ymd("2016/09/13") ## Just figure it out
mdy("09-13-2016") ## Mostly U.S.
dmy("13-09-2016") ## Europe
```
Even if the individual dates are formatted differently, ymd() can usually figure it out.
```{r}
x <- c(
    "2016-04-05",
    "2016/05/06",
    "2016,10,4"
)
ymd(x)

```
## 2. From individual date-time components

```{r multiple columns}
install.packages("nycflights13")
library("nycflights13")
library(tidyverse)
flights %>%
    select(year, month, day)

```

To create a date/time from this sort of input, use

*make_date(year, month, day)* for dates, or
*make_datetime(year, month, day, hour, min, sec, tz)* for date-times
We combine these functions inside of mutate to add a new column to our dataset:
```{r combine}
flights %>%
    select(year, month, day) %>%
    mutate(departure = make_date(year, month, day))

# The flights also contains a hour and minute column.
x <- flights %>%
    select(year, month, day, hour, minute) %>%
  mutate(departure_time = make_datetime(year,month, day,hour = hour, min = minute))

x$departure_time[1] # default is UTC
```

## 3. From other types
You may want to switch between a date-time and a date.

That is the job of *as_datetime()* and *as_date()*
```{r switch}
today()
as_datetime(today())
now()
as_date(now())
```
## Date-Times in R
### From a string
ymd() and friends create dates.

To create a date-time from a character string, add an underscore and one or more of “h”, “m”, and “s” to the name of the parsing function:

Times can be coerced from a character string with ymd_hms()
```{r}
ymd_hms("2017-01-31 20:11:59")
mdy_hm("01/31/2017 08:01")

# You can also force the creation of a date-time from a date by supplying a timezone:
ymd_hms("2016-09-13 14:00:00")
ymd_hms("2016-09-13 14:00:00", tz = "America/New_York")
ymd_hms("2016-09-13 14:00:00", tz = "")
```
## POSIXct
the POSIXct class represents the number of seconds since 1 January 1970. (
c "calendar time"
 'POSIXct' stands for “Portable Operating System Interface”, calendar time.)
```{r}
 x <- ymd_hm("1970-01-01 01:00")
class(x)
unclass(x)
typeof(x)
attributes(x)
```
## POSIXlt
POSIXlt is a list underneath and it stores a bunch of other useful information like the day of the week, day of the year, month, day of the month
```{r}
y <- as.POSIXlt(x)
y
typeof(y)
attributes(y)

```

POSIXlts are rare inside the tidyverse

POSIXct’s are always easier to work with, so if you find you have a POSIXlt, you should always convert it to a regular data time lubridate::as_datetime().


## Time Zones!
Time zones were created to make your data analyses more difficult as a data analyst.
*ymd_hms()* function will by default use Coordinated *Universal Time (UTC)* as the time zone
Specifying tz = "" in one of the ymd() and friends functions will use the local time zone
```{r}
x <- ymd_hm("1970-01-01 01:00", tz = "")
x
attributes(x)

# The tzone attribute is optional. It controls how the time is printed, not what absolute time it refers to.
attr(x, "tzone") <- "US/Pacific"
x
attr(x, "tzone") <- "US/Eastern"
x
```
### Operations on Dates and Times
Arithmetic
You can add and subtract dates and times.
```{r}
x <- ymd("2012-01-01", tz = "") ## Midnight
y <- dmy_hms("9 Jan 2011 11:34:21", tz = "")
x - y ## this works
```
You can do comparisons too (i.e. >, <, and ==)
```{r compare}
x < y ## this works
x > y ## this works
x + y ## what??? why does this not work?
# The class of x is POSIXct.
class(x)
```
POSIXct objects are a measure of seconds from an origin, usually the UNIX epoch (1st Jan 1970).

Just add the requisite number of *seconds* to the object:
```{r}
x + 3 * 60 * 60 # add 3 hours
```
Same goes for days. For example, you can just keep the date portion using date():
```{r}
y <- date(y)
y
# And then add a number to the date (in this case 1 day)
y + 1
```

## Leaps and Bounds
Even keeps track of leap years, leap seconds, daylight savings, and time zones.

Leap years
```{r leap year}
x <- ymd("2012-03-01")
y <- ymd("2012-02-28")
x - y
```
Not a leap year
```{r not leap year}
x <- ymd("2013-03-01")
y <- ymd("2013-02-28")
x - y
```
BUT beware of time zones!
```{r leap seconds}
x <- ymd_hms("2012-10-25 01:00:00", tz = "")
y <- ymd_hms("2012-10-25 05:00:00", tz = "GMT")
y - x
.leap.seconds
```


## Extracting Elements of Dates/Times
There are a set of helper functions in lubridate that can extract sub-elements of dates/times
```{r}
x <- ymd_hms(c(
    "2012-10-25 01:13:46",
    "2015-04-23 15:11:23"
), tz = "")
year(x)
day(x)
weekdays(x)
x <- month(x) +1
x
```

## Time Elements

## Visualizing dates
```{r}
library(here)
library(readr)
storm <- read_csv(here("data", "storms_2004.csv.gz"), progress = FALSE)
storm
names(storm)
```

Takss:
Create a subset of the storm dataset with only the four columns above.
Create a new column called begin that contains the BEGIN_DATE_TIME that has been converted to a date/time R object.
Rename the EVENT_TYPE column as type.
Rename the DEATHS_DIRECT column as deaths.
```{r}
storm %>%
  select( BEGIN_DATE_TIME, EVENT_TYPE, DEATHS_DIRECT) %>%
  mutate(begin = dmy_hms(BEGIN_DATE_TIME)) %>%
  rename(type = EVENT_TYPE,deaths = DEATHS_DIRECT ) 


```

## Histograms of Dates/Times
We can make a histogram of the dates/times to get a sense of when storm events occur.
```{r}
library("ggplot2")
storm_sub %>%
    ggplot(aes(x = begin)) +
    geom_histogram(bins = 20) +
    theme_bw()
```
We can group by event type too.
```{r}
library(ggplot2)
storm_sub %>%
    ggplot(aes(x = begin)) +
    facet_wrap(~type) +
    geom_histogram(bins = 20) +
    theme_bw() +
    theme(axis.text.x.bottom = element_text(angle = 90))
```

### Scatterplots of Dates/Times
focus on a single month, single day
```{R}
storm_sub %>%
    filter(month(begin) == 6, day(begin) == 16) %>%
    ggplot(aes(begin, deaths)) +
    geom_point()
```

## Final question
1. What happens if you parse a string that contains invalid dates?

```{r}
ymd(c("2010-10-10", "bananas"))  # default give a warning and NA

## Compare against base R's behavior:
as.Date(c("2010-10-10", "bananas"))  # don't say anything and you wouldn't know anything

```
2. What does the tzone argument to today() do? Why is it important?
unclass(today())
```{r}
unclass(today())
```
3. Use the appropriate lubridate function to parse each of the following dates:
```{r}
d1 <- "January 1, 2010"
d2 <- "2015-Mar-07"
d3 <- "06-Jun-2017"
d4 <- c("August 19 (2015)", "July 1 (2015)")
d5 <- "12/30/14" # Dec 30, 20

d1 <- mdy(d1)
d1
```



#
# 21-Regular experssion 
# regex basics
1. searching for a pattern or string within another string (e.g searching for the string “a” in the string “Maryland”)
2. replacing one part of a string with another string (e.g replacing the string “t” with “p” in the string “hot” where you are changing the string “hot” to “hop”)

## string basics
In R, you can create (character) strings with either single quotes ('hello!') or double quotes ("hello!") – no difference (not true for other languages!).

I recommend using the double quotes, 
```{r}
string1 <- "This is Siyu's string"
string2 <- 'If I want to include a "quote" inside a string, I use single quotes'
```

### grepl()
One of the most basic functions in R that uses regular expressions is the *grepl(pattern, x)* function, which takes *two arguments* and *returns a logical*:

1. A regular expression (pattern)
2. A string to be searched (x)
If the string (x) contains the specified regular expression (pattern), then grepl() will return TRUE,otherwise FALSE
```{r grepl}
regular_expression <- "a"
string_to_search <- "Maryland"

grepl(pattern = regular_expression, x = string_to_search)

grepl("land", "Maryland")
grepl("Marly", "Maryland")
```
#### example
```{r}
head(state.name)
length(state.name)
```
#### metacharacters
The first metacharacter that we will discuss is "."

The metacharacter that only consists of a period represents any character other than a *new line* (we will discuss new lines soon).
```{r metacharacters}
grepl(".", "Maryland")
grepl(".", "*&2[0+,%<@#~|}")
grepl(".", "")
```
As you can see the period metacharacter is very liberal.

This metacharacter is most useful when you *do not care about a set of characters* in a regular expression.
```{r metacharacters 2}
grepl("a.b", c("aaa", "aab", "abb", "acadb"))
```
In the case above, grepl() returns TRUE for all strings that contain an a followed by any other character followed by a b.


## repetition
You can specify a regular expression that contains a certain number of characters or metacharacters using the enumeration metacharacters (or sometimes called *quantifiers*).

1. +: indicates that one or more of the preceding expression should be present (or matches at least 1 time)
2. *: indicates that zero or more of the preceding expression is present (or matches at least 0 times)
3. ?: indicates that zero or 1 of the preceding expression is not present or present at most 1 time (or matches between 0 and 1 times)

### example
```{r repetition example}
args(grepl)
# Does "Maryland" contain one or more of "a" ?
grepl("a+", "Maryland")
# Does "Maryland" contain one or more of "x" ?
grepl("x+", "Maryland")
# Does "Maryland" contain zero or more of "x" ?
grepl("x*", "Maryland")  #  TRUE
grepl("Lx*", "Maryland") # FALSE
```

If you want to do more than one character, you need to wrap it in ().
```{r wrap}
# Does "Maryland" contain zero or more of "x" ?
grepl("(xx)*", "Maryland")
grepl("La(la)*", "Maryland")


grepl("(zz)+", "spookyhalloween") # Does “zz” appear 1 or more times?
grepl("(ee)+", "spookyhalloween")
grepl("ee+", "spookyhalloween")   # I want find the e second times or more
grepl("(oo)*", "spookyhalloween") # Does “oo” appear 0 or more times?
grepl("zz+", "zaz")   # FALSE 
grepl("(zz)+", "zaz")   # FALSE 

```

You can also specify exact numbers of expressions using curly brackets {}.
{n}: exactly n
{n,}: n or more
{,m}: at most m
{n,m}: between n and m
For example "a{5}" specifies “a exactly five times”, "a{2,5}" specifies “a between 2 and 5 times,” and "a{2,}" specifies “a at least 2 times.” Let’s take a look at some examples:
```{r}
# Does "Mississippi" contain exactly 2 adjacent "s" ?
grepl("s{2}", "Mississippi")
# This is equivalent to the expression above:
grepl("ss", "Mississippi")
# Does "Mississippi" contain between 1 and 3 adjacent "s" ?
grepl("s{1,3}", "Mississippi")
# Does "Mississippi" contain between 2 adjacent "ss" ?
grepl("(ss){2}", "Mississippi") # FALSE means to find "ssss" 
grepl("(ss)+", "Mississippi") 
grepl("i.{2}i.{2}.{2}", "Mississippi") 

# Does "Mississippi" contain between 2 and 4 adjacent "i" ?
grepl("i{2,4}", "Mississippi")
# Does "Mississippi" contain between 2 adjacent "iss" ?
grepl("(iss){2}", "Mississippi")

# Does "Mississippi" contain the pattern of an "i" followed by
# 2 of any character, with that pattern repeated three times adjacently?
grepl("(i.{2}){3}", "Mississippi")   
```

Let’s practice a few out together. Make the following regular expressions for the character string “spookyspookyhalloweenspookyspookyhalloween”:

Search for “spooky” exactly 2 times. What about 3 times?
Search for “spooky” exactly 2 times followed by any character of length 9 (i.e. “halloween”).
Same search as above, but search for that twice in a row.
Same search as above, but search for that three times in a row.

 grepl can't tell us the specific times, it just tell us yes or no
```{r}
grepl("(spooky){2}", "spookyspookyhalloweenspookyspookyhalloween")
grepl("(spooky){2}.{9}", "spookyspookyhalloweenspookyspookyhalloween")

grepl("((spooky){2}.{9}){2}", "spookyspookyhalloweenspookyspookyhalloween")
grepl("((spooky){2}.{9}){3}", "spookyspookyhalloweenspookyspookyhalloween")

```



## capture group
In the examples above, I used parentheses () to create a capturing group. 
In the “Mississippi” example, I first created the regex "i.{2}" which matches i followed by any two characters (“iss” or “ipp”). Then, I used a capture group to wrap that regex, and to specify exactly three adjacent occurrences of that regex.
You can specify sets of characters (or character sets or character classes) with regular expressions, some of which come built in, but you can build your own character sets too.

### character sets
First, we will discuss the built in character sets:

* words ("\\w") = Words specify any letter, digit, or a underscore
* digits ("\\d") = Digits specify the digits 0 through 9
* whitespace characters ("\\s") = Whitespace specifies line breaks, tabs, or spaces
Each of these character sets have their own compliments:
* not words ("\\W")
* not digits ("\\D")
* not whitespace characters ("\\S")

"\n": newline
"\t": tab,
```{r}
"\\d"
double_quote <- "\""
double_quote
single_quote <- "'"
single_quote
```
In fact, putting two backslashes before any punctuation mark that is also a metacharacter indicates that you are looking for the symbol and not the metacharacter meaning.

For example "\\." indicates you are trying to match a period in a string. Let’s take a look at a few examples:
```{r}
grepl("\\+", "tragedy + time = humor")
grepl("\\.", "https://publichealth.jhu.edu")
```

You will also sometimes see strings like “0b5”, this is a way of writing non-English characters that works on all platforms:
```{r}
x <- c("\\t", "\\n", "\u00b5")
x
```

## brackets
You can also specify specific character sets using straight brackets [].

For example a character set of just the vowels would look like: "[aeiou]".

```{r brackets}
grepl("[aeiou]", "rhythms")
```

You can find the complement to a specific character by putting a carrot ^ after the first bracket. 
For example "[^aeiou]" matches all characters except the lowercase vowels.

```{r}
grepl("[^aeiou]", "rhythms")
```

## ranges
You can also specify ranges of characters using a hyphen - inside of the brackets.

For example:

"[a-m]" matches all of the lowercase characters between a and m
"[5-8]" matches any digit between 5 and 8 inclusive
```{r}
grepl("[a-m]", "xyz")  # [1] FALSE
grepl("[a-m]", "ABC")  # [1] FALSE
grepl("[a-m]", tolower("ABC"))  # [1] TRUE
grepl("[a-mA-M]", "ABC")
```
## beginning and end
There are also metacharacters for matching the beginning and the end of a string which are "^" and "$" respectively.

Let’s take a look at a few examples:
```{r}
grepl("^a", c("bab", "aab")) # [1] FALSE  TRUE
grepl("b$", c("bab", "aab", "b")) # [1] TRUE TRUE TRUE
grepl("^[ab]*$", c("bab", "aab", "abc")) # [1]  TRUE  TRUE FALSE

grepl("[ab]$", c("bab", "aab", "abc")) # [1]  TRUE  TRUE FALSE
```
## OR metacharacter
The last metacharacter we will discuss is the OR metacharacter ("|").

The OR metacharacter matches either the regex on the left or the regex on the right side of this character. A few examples:
```{r}
grepl("a|b", c("abc", "bcd", "cde"))
```



# Other regex in base R
First, we will review our workhorse of this lesson, grepl(), which stands for “grep logical.”

grepl("[Ii]", c("Hawaii", "Illinois", "Kentucky"))

## grep()
Then, there is old fashioned grep(pattern, x), which returns the indices of the vector that match the regex:

grep(pattern = "[Ii]", x = c("Hawaii", "Illinois", "Kentucky"))


## sub()
The sub(pattern, replacement, x) function takes as arguments a regex, a “replacement,” and a vector of strings. This function will replace the first instance of that regex found in each string.

sub(pattern = "[Ii]", replacement = "1", x = c("Hawaii", "Illinois", "Kentucky"))

## gsub()
The gsub(pattern, replacement, x) function is nearly the same as sub() except it will replace every instance of the regex that is matched in each string.


## strsplit()
The strsplit(x, split) function will split up strings (split) according to the provided regex (x) .

If strsplit() is provided with a vector of strings it will return a list of string vectors.
```{r}
two_s <- state.name[grep("ss", state.name)]
two_s
strsplit(x = two_s, split = "ss")
strsplit(x= ("2023-10-05_class.R"), split = "_" )
```



